### 2012-07-31 paper/paper.tex

%\documentclass[useAMS,usenatbib]{biom}
\documentclass[useAMS,usenatbib,referee]{example/biom}

%
%
%  Papers submitted to Biometrics should ALWAYS be prepared
%  using the referee option!!!!

%%%%% AUTHORS - PLACE YOUR OWN MACROS HERE %%%%%

\def\bSig\mathbf{\Sigma}
\newcommand{\VS}{V\&S}
\newcommand{\tr}{\mbox{tr}}

%  If you have a landscape table you need to use the rotating package

\usepackage[figuresright]{rotating}
\graphicspath{{example/}{graphics/}}
%% \raggedbottom % To avoid glue in typesetteing, sbs>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\setcounter{footnote}{2}

\title[This is an Example of Recto Running Head]{This is an Example of
an Article Title This is an Example of an Article Title}

\author{B. Calderhead$^{*}$\email{email@address.com} \\
	   Institute of Physics, CA 560034, USA
	   \and 
	   C. S. Gillespie$^{*}$\email{colin.gillespie@newcastle.ac.uk}\\
	   School of Mathematics \& Statistics, Newcastle University,
	   NE1 7RU, UK
	   }

\begin{document}

\date{}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2012}

\volume{XX}
\artmonth{YYYY}
\doi{10.1111/j.1541-0420.2005.00454.x}

%  This label and the label ``lastpage'' are used by the \pagerange
%  command above to give the page range for the article

\label{firstpage}

%  pub the summary here

\begin{abstract}
The world will little note, nor long remember, what we say here, but
can never forget what they did here. It is for us, the living, rather
to be dedicated here to the unfinished work which they have, thus far,
so nobly carried out. It is rather for us to be here dedicated to the
great task remaining before us--that from these honored dead we take
increased devotion to that cause for which they here gave the last
full measures of devotion--that we here highly resolve that these dead
shall not have died in vain; that this nation shall have a new birth
of freedom; and that this government of the people, by the people, for
the people, shall not perish from the earth. The world will little
note.
\end{abstract}

%
%  Please place your key words in alphabetical order, separated
%  by semicolons, with the first letter of the first word capitalized,
%  and a period at the end of the list.
%

\begin{keywords}
Stochastic; Moment-closure.
\end{keywords}

\maketitle

\section{Introduction}
\label{s:intro}

\section[]{Material Description of the Envelope\break Predominantly Model}

\subsection{Radial Distribution of Dust}

\subsection{An Example of Head two}
\subsubsection{An example of head three comparison with oxygen and~carbon~Miras}

% \begin{figure*}
% %  \centerline{\includegraphics[width=4.5in]{mouse.eps}}
%   \caption{An example of figure caption. An example of 
%   figure caption. An example of figure caption. An example of
%    figure caption. An example of figure caption.}
% \label{f:bigmouse}
% \end{figure*}

\section{Discussion}

\backmatter

%%%%%% include this section if you wish to acknowledge people,
%%%%%% grant support, etc.

\section*{Acknowledgements}

The authors thank Professor A. Sen for some helpful suggestions,
Dr C. R. Rangarajan for a critical reading of the original version of the
paper, and an anonymous referee for very useful comments that improved
the presentation of the paper.\vspace*{-8pt}

%%%%%% include this section only if your manuscript refers to supplementary
%%%%%% materials -- see Instructions for Authors at 
%%%%%% http://www.tibs.org/biometrics

\section*{Supplementary Materials}

Web Appendix 1 referenced in Section~\ref{ss:example} is available
with this paper at the Biometrics website on Wiley Online Library.
\vspace*{-8pt}

\begin{thebibliography}{}
\bibitem[\protect\citeauthoryear{Akash and Tirky}{1988}]{b24} 
Akash, F. J. and Tirky, T. (1988). Proper multivariate conditional
autoregressive models for spatial data analysis. {\it Biometrics} {\bf 196,} 173.

\bibitem[\protect\citeauthoryear{Balaram et~al.}{1985a}]{b2} 
Balaram, C. A., Neeraj, G., Haq, H. J., and Cliff, P. E., (1985a). {\it
NIFT\/} 2nd edition. Boca Raton, Florida: Chapman and Hall.

\bibitem[\protect\citeauthoryear{Commel}{1961}]{b18} 
Commel, J. K. (1961). {\it National Sangget Academia}, 2nd edition.
Hoboken, New Jersey: Wiley.

\bibitem[\protect\citeauthoryear{Das  and Patra}{1986}]{b25} 
Das, B. and Patra, H. M. (1986). Using counts to simultaneously estimate
abundance and detection probabilities in a salamander community. {\it
Herpetologica} {\bf 60,} 468--478.

\bibitem[\protect\citeauthoryear{Emrat}{1974}]{b14} Emrat, T. (1974).
{\it Topics in Stochastic Processes.} New York: Academic  Press.

\bibitem[\protect\citeauthoryear{Emrat}{1985}]{b15} Emrat, T. (1985).
{\it Topics in Stochastic Processes.} New York: Academic  Press.

\bibitem[\protect\citeauthoryear{Goldman et~al.}{1987}]{b9} 
Goldman, M. J. and Ewin, A. (1987). A comparison of smoothing
techniques for CD4 data measured with error in a time-dependent Cox
proportional hazards model. {\it Statistics in Medicine} {\bf 17,} 2061--2077.

\bibitem[\protect\citeauthoryear{Goodsman}{1972}]{b8} 
Goodsman, R. C. (1972). Testing hypotheses in the functional linear
model. {\it Scandinavian Journal of Statistics} {\bf 30,} 241--251.

\bibitem[\protect\citeauthoryear{Gooms}{1972}]{b5} Gooms, R. D.
(1972). Goodsman, R.C. (1972). Testing hypotheses in the functional linear
model. {\it Scandinavian Journal of Statistics} {\bf 31,} 315--323.

\bibitem[\protect\citeauthoryear{Gooms  and Wool}{1970}]{b7} Gooms,
R. D. and Wool, N. J. (1970). A comparison of smoothing
techniques for CD4 data measured with errors in a time-dependent Cox
proportional hazards model. {\it Statistics in Medicine} {\bf 17,} 2091--2099.

\bibitem[\protect\citeauthoryear{Hackel}{1985}]{b10} 
Hackel, P. (1985). {\it Topics in Stochastic Processes.} New York: Academic  Press.

\bibitem[\protect\citeauthoryear{\nobreak Holloman, \nobreak Tinku,  and Gerg}{Holloman et~al.}{1979}]{b11}
Holloman, P. M., Tinku, H. A., and Gerg, I. (1979). A comparison of smoothing
techniques for CD4 data measured with techniques a time-dependent Cox
proportional hazards model. {\it Statistics in Medicine} {\bf 17,} 2105--2111.

\bibitem[\protect\citeauthoryear{Juman}{1986}]{b12} 
Juman, M. (1986). Using counts to simultaneously estimate
abundance and detection probabilities in a salamander community. {\it
Herpetologica} {\bf 61,} 482--495.

\bibitem[\protect\citeauthoryear{Kartik B}{1969}]{b13} 
Kartik, B. V. (1969). Using counts to simultaneously estimate
abundance and detection probabilities in a salamander community. {\it
Herpetologica} {\bf 62,} 511--519.

\bibitem[\protect\citeauthoryear{Lenin}{1932}]{b17} 
Lenin, D. B. (1932). A comparison of smoothing
techniques for CD4 data measured with errors in a time-dependent Cox
proportional hazards model. {\it Statistics in Medicine} {\bf 17,} 2125--2135.

\bibitem[\protect\citeauthoryear{O'Rourke}{1979}]{b4} O'Rourke, D. (1996). Industrial ecology: A 
critical review. {\it International Journal of Environment and Pollution}
{\bf 6,} 389--112.

\bibitem[\protect\citeauthoryear{Oman and \nobreak Raj}{1986}]{b19}
Oman, F. M. and  Raj, E. (1986). {\it Topics in Stochastic Process.} New York: Academic  Press.

\bibitem[\protect\citeauthoryear{Prem et~al.}{1963}]{b20} Prem, G. W., Christ, W., Smap, J.,
Wills, J. A. (1963). {\it Topics in Stochastic Processes.} New York: Academic  Press.

\bibitem[\protect\citeauthoryear{Rahul}{1981}]{b1} Rahul, S. R. (1981).
Industrial ecology: A  critical review. {\it International Journal of
Environment and Pollution} {\bf 6,} 371--375.

\bibitem[\protect\citeauthoryear{Ram  and Shyam}{1972}]{b6} Ram, R. D.
and Shyam, S. (1972). {\it Topics in Stochastic Process.} New York: Academic  Press.

\bibitem[\protect\citeauthoryear{Ravan  and Hari}{1983a}]{b21} Ravan,
M. and Hari, S. (1983). Industrial ecology: A  critical review.
{\it International Journal of Environment and Pollution} {\bf 6,} 389--395.

\bibitem[\protect\citeauthoryear{Ravan  and Hari}{1983b}]{b22}
Ravan, M. and Hari, S. (1984). Industrial ecology: A  critical
review. {\it International Journal of Environment and Pollution} {\bf 7,} 512--519.

\bibitem[\protect\citeauthoryear{Sharma et~al.}{1984}]{b16} Sharma,
F. J. (1984). {\it Topics in Stochastic Process.} New York: Academic Press.

\bibitem[\protect\citeauthoryear{Vim   and Potter}{1988}]{b23} Vim, W.
E. and Potter, H. J. (1988). A  critical
review. {\it International Journal of Environment and Pollution} {\bf 9,} 635--641.

\end{thebibliography}

\label{lastpage}

\end{document}